Kindly Bent to Free Us

### 2019-02-27 kindly/infer/infer.tex
Smooth over the borrow-related bits.
snapshot
Add checking during normalization.

\begin{figure}[tp]\centering
\begin{mathpar}
  % \Sv\cap\Sv'
  % &= \left\{ x \to l,l'
  %   \mid \Sone{x}{l}\in\Sv \text{ and } \Sone{x}{l'}\in\Sv' \right\} \\
  \infer
  { \forall x \in \Dom{\Sv_1}\cup\Dom{\Sv_2}.\quad
    \tsplit{C_x}{\Sv(x)}{\Sv_1(x)}{\Sv_2(x)}
  } 
  { \tsplit{\bigwedge C_x}{\Sv}{\Sv_1}{\Sv_2} }
  % &= \Sv\cap\Sv' \sqcup
  %   (\Sv\Sdel{x' \mid x'\in\Sv'}) \sqcup
  %   (\Sv'\Sdel{x \mid x\in\Sv})
  %   &\text{where $\sqcup$ is the disjoint union}\\
  \and
  \inferrule{}
  { \Cleq{\Sv}{k_0}
    = \bigwedge_{\Sone{x}{l_x}\in\Sv} \bigwedge_{k\in l_x} \Cleq{k}{k_0}
  }
\end{mathpar}
\caption{Operation on usage maps $\Sv$}
\label{op:usgmap}
\end{figure}

@@

\begin{figure}[h]
  \centering
  \input{infer/kinds}
  \caption{Kind inference rules -- $\inferK{(C,\unif)}{\E}{\tau}{k}$}
  \label{rules:kinding}
\end{figure}

@@

\begin{figure*}[h]
  \input{infer/typing}
  \caption{Type Inference rules -- $\inferW{\Sigma}{(C,\psi)}{\bf{\E}}{\bf{e}}{\tau}$ }
  \label{rules:typing}
\end{figure*}

\subsection{Normal forms and generalization}
\label{sec:normalize}

@@

  We also complement $\mathcal R$ with $(\leq)$ by adding an edge
  between related constants.
\item For each strongly connected component in $\mathcal G$, unify all its vertices and replace it by a representative.
  We note $\unif_k$ the substitution that replaces a kind variable by
  its representative.
  The representative of a strongly connected component $g$ can be determined as follows:
  \begin{itemize}
  \item If $g$ does not contain any constant, then the representative
    is a fresh kind variable.
  \item If $g$ contains exactly one constant, it is the representative.
  \item Otherwise, the initial constraint $C_0$ is not satisfiable.
  \end{itemize}
  Note that this step will also detect all unsatisfiable constraints.
\item Take the transitive closure of $\mathcal R$.
\item Remove all the vertices corresponding to the kind variables $\kvar_i$
  that are existentially quantified in $C^k$.
\item Take the transitive reduction of $\mathcal R$.
\item Remove the extremums of $\mathcal L$ and the edges of $(\leq)$
  from $\mathcal R$.
\item Return $C = \left\{ k \leq k' \mid k \operatorname{\mathcal R}k' \right\}$
  and $\unif =  \unif_\tau \meet \unif_k$.
\end{enumerate}

@@

  Let $C^{A}$ and $\unif_k$ be the results after step (3).
  By definition, we have $\unif_k C^{k+} \equivC C^{A}$ and
  $\unif_k C^{A} = C^{A}$. Since $\phi_k$ has already be applied to $C$ before
  unifying the strongly connected components,
  we have that $\phi_k\leq\unif_k$.

@@

  Steps (4) to (8) all preserve the free variables and the equivalence
  of constraints, which concludes.
\end{proof}

@@

  Let us note $C_1'$ and $C_2'$ the constraints after step (4). $P(C_1')$ and
  $P(C_2')$ hold. Indeed, since $C_1'$ and $C_2'$ are only composed
  of existential quantifications and kind inequalities, the only rules
  that applies are transitivity and lattice inequalities.
  After step (2) and (4), the associated relations are fully saturated for these
  two rules, hence all inequalities that can be deduced from $C_a'$ are already
  present in the relation.

  The property $P$ is preserved by step (5) since we only remove
  inequalities that involve existentially quantified variables. Such
  inequalities could not be picked in $P$.
  % In (6) we only remove
  % trivial lattice inequalities.

  Let us note $C''_a$ for $a\in\{1,2\}$ the constraints after step (5).
  Since there are no more existential variables,
  we have $C''_a=(k_i,k'_i)_i=\mathcal R''_a$.
  For any $C=(k,k')$ such that $\entail{}{C}$ and $\entail{C''_a}{C}$,
  then $C \in (\leq) \subset {\mathcal R}''_a$.
  Indeed, the only trivial inequalities in our system are equalities of the form
  $(\kvar, \kvar)$, which were removed in step (3) and the lattice inequalities.

@@

  This equality is preserved by step (6) and (7)
  since the transitive reduction of a directed acyclic graph is unique,
  which concludes.
\end{proof}

>>>>

\begin{figure}[tp]\centering
\begin{mathpar}
  % \Sv\cap\Sv'
  % &= \left\{ x \to l,l'
  %   \mid \Sone{x}{l}\in\Sv \text{ and } \Sone{x}{l'}\in\Sv' \right\} \\
  \infer
  { \forall x \in \Dom{\Sv_1}\cup\Dom{\Sv_2}.\quad
    \tsplit{C_x}{\Sv(x)}{\Sv_1(x)}{\Sv_2(x)}
  } 
  { \tsplit{(\bigwedge C_x)}{\Sv}{\Sv_1}{\Sv_2} }
  % &= \Sv\cap\Sv' \sqcup
  %   (\Sv\Sdel{x' \mid x'\in\Sv'}) \sqcup
  %   (\Sv'\Sdel{x \mid x\in\Sv})
  %   &\text{where $\sqcup$ is the disjoint union}\\
  \and
  \inferrule{}
  { \Cleq{\Sv}{k_0}
    = \bigwedge_{\Sone{x}{l_x}\in\Sv} \bigwedge_{k\in l_x} \Cleq{k}{k_0}
  }
  \and
  \inferrule
  { \forall \bshadow{x} \in \Sv, b'\in\Sv'.\quad
    \tregion{C_x}{y}{\bshadow{x}}{b'} }
  { \tregion{(\bigwedge C_x)}{y}{\Sv}{\Sv'} }
  \and
  % \inferrule
  % { \text{if } \operatorname{lenght}(\Sv(x))  = \Cempty{} \wedge  }
  { \Weaken_{(x,k)}(\Sv) =
    \begin{dcases}
      \Ctrue& \operatorname{length}(\Sv(x)) = 1\\
      {k}\lk\kaff_\infty& \text{ otherwise}\\
    \end{dcases}

  }
\end{mathpar}
\caption{Operation on usage maps $\Sv$}
\label{op:usgmap}
\end{figure}

@@

\begin{figure}[ht]
  \centering
  \input{infer/kinds}
  \caption{Kind inference algorithm -- $\inferK{(C,\unif)}{\E}{\tau}{k}$}
  \label{rules:kinding}
\end{figure}

\begin{figure}[ht]
  \centering
  \input{infer/sd-kinds}
  \caption{Syntax-directed kind inference -- $\inferSK{C}{\E}{\tau}{k}$}
  \label{rules:sd-kinding}
\end{figure}

@@

\begin{figure*}[h]
  \input{infer/typing}
  \caption{Type inference algorithm -- $\inferW{\Sigma}{(C,\psi)}{\bf{\E}}{\bf{e}}{\tau}$ }
  \label{rules:typing}
\end{figure*}

\begin{figure*}[ht]
  \input{infer/sd-typing}
  \caption{Syntax-directed typing rules}
  \label{fig:syntax-directed-typing}
\end{figure*}
\subsection{Normal forms and generalization}
\label{sec:normalize}

@@

  We also complement $\mathcal R$ with $(\leq)$ by adding an edge
  between related constants.
\item
  At this point, we can easily check for satisfiability: A constraint
  is satisfiable (in the given environment) if and only if,
  for any constants $l_1$ and $l_2$ such that
  there is a path from $l_1$ to $l_2$ in $\mathcal G$, then $l_1\lk_\Lat l_2$.
  If this is not the case, we return \textbf{fail}.

\item For each strongly connected component in $\mathcal G$, unify all its vertices and replace it by a representative.
  We note $\unif_k$ the substitution that replaces a kind variable by
  its representative.
  The representative of a strongly connected component $g$ can be determined as follows:
  \begin{itemize}
  \item If $g$ does not contain any constant, then the representative
    is a fresh kind variable.
  \item If $g$ contains exactly one constant, it is the representative.
  \item Otherwise, the initial constraint $C_0$ is not satisfiable.
  \end{itemize}
  Note that this step will also detect all unsatisfiable constraints.
\item Take the transitive closure of $\mathcal R$.
\item Remove all the vertices corresponding to the kind variables $\kvar_i$
  that are existentially quantified in $C^k$.
\item Take the transitive reduction of $\mathcal R$.
\item Remove the extremums of $\mathcal L$ and the edges of $(\leq)$
  from $\mathcal R$.
\item Return $C = \left\{ k \leq k' \mid k \operatorname{\mathcal R}k' \right\}$
  and $\unif =  \unif_\tau \meet \unif_k$.
\end{enumerate}

@@

  Let $C^{A}$ and $\unif_k$ be the results after step (4).
  By definition, we have $\unif_k C^{k+} \equivC C^{A}$ and
  $\unif_k C^{A} = C^{A}$. Since $\phi_k$ has already be applied to $C$ before
  unifying the strongly connected components,
  we have that $\phi_k\leq\unif_k$.

@@

  Steps (5) to (9) all preserve the free variables and the equivalence
  of constraints, which concludes.
\end{proof}

@@

  Let us note $C_1'$ and $C_2'$ the constraints after step (4). $P(C_1')$ and
  $P(C_2')$ hold. Indeed, since $C_1'$ and $C_2'$ are only composed
  of existential quantifications and kind inequalities, the only rules
  that applies are transitivity and lattice inequalities.
  After step (2) and (5), the associated relations are fully saturated for these
  two rules, hence all inequalities that can be deduced from $C_a'$ are already
  present in the relation.

  The property $P$ is preserved by step (6) since we only remove
  inequalities that involve existentially quantified variables. Such
  inequalities could not be picked in $P$.
  % In (6) we only remove
  % trivial lattice inequalities.

  Let us note $C''_a$ for $a\in\{1,2\}$ the constraints after step (5).
  Since there are no more existential variables,
  we have $C''_a=(k_i,k'_i)_i=\mathcal R''_a$.
  For any $C=(k,k')$ such that $\entail{}{C}$ and $\entail{C''_a}{C}$,
  then $C \in (\leq) \subset {\mathcal R}''_a$.
  Indeed, the only trivial inequalities in our system are equalities of the form
  $(\kvar, \kvar)$, which were removed in step (4) and the lattice inequalities.

@@

  This equality is preserved by step (7) and (8)
  since the transitive reduction of a directed acyclic graph is unique,
  which concludes.
\end{proof}
